---
title: "Beta RGC mosaics"
author: "Stephen J Eglen"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r rsetup-knitr,eval=TRUE,include=FALSE}
require(knitr)
options(width=60)
opts_chunk$set(cache=TRUE)
```


# Initialisation

We depend on several packages from CRAN, and several written by SJE:

```{r}
library(sjedmin)
library(sjedist)
library(spatstat)
```

## Load the data (observed mosaics and parameter estimates)

```{r}
attach("bivariate_mosaics.Rda")
## Load in the parametric fits for univariate mosaics.
load("rgc_parests.Rda")
```

## Function definitions

```{r}
## First set up h for all sims.
hpar <- function(d,theta) {
  ## Choice of h() suggested by Peter.
  delta<-theta[1]
  phi<-theta[2]
  alpha<-theta[3]
  res <- (0*(d<delta)) + (d>=delta)*(1-exp(-((d-delta)/phi)^alpha))
  if (any (is.nan(res)))
    res[ which(is.nan(res))] <- 0

  res
}

ranking <- function(arr) {
  ## Evaluate fit of row 1 (the real data) with remaining rows (simulations)
  ## using equations 6.3--6.5 from (Diggle, 2002, page 89).
  n.s <- nrow(arr)
  u <- rep(0, n.s)
  for (i in 1:n.s) {
    ave.i <- apply( arr[-i,], 2, sum) / (n.s - 1)
    u[i] <- sum((ave.i - arr[i,])^2)
    ##u[i] <- max( abs( ave.i - arr[i,]) )
  }
  signif( (rank(-u)[1])/n.s, 5)
}

ranking.k <- function(arr) {
  ## Assume we have a K function; cf it against the pi t^2 version.
  n.s <- nrow(arr)
  u <- rep(0, n.s)
  for (i in 1:n.s) {
    ave.i <- apply( arr[-i,], 2, sum) / (n.s - 1)
    u[i] <- sum((ave.i - arr[i,])^2)
    ##u[i] <- max( abs( ave.i - arr[i,]) )
  }
  signif( (rank(-u)[1])/n.s, 5)
}


make.ci <- function(da, fns=c("l1", "f1")) {
  ## Make the confidence intervals for each of the arrays mentioned in fns.
  ## Return these confidence intervals as a list.

  res <- list()
  for (f in fns) {
    arr <- da$get()[[f]]
    p <- 0.05                             # 5% for 95% confidence intervals 
    ci <- apply(arr, 2, quantile, probs=c(p/2, 1-(p/2)))
    
    ## This would be nice, but doesn't work.
    ## attr(dist.arr$get()[[f]], "ci") <- ci
    
    new <- list(ci); names(new) <- f
    res <- c(res, new)
    
    res
  }
}


kranking <- function(arr) {
  ## Assume we have a L function; cf it against the pi t^2 version.
  n.s <- nrow(arr)
  u <- rep(0, n.s)

  t <- as.real(colnames(arr))
  theor <- sqrt(pi *t^2)
  for (i in 1:n.s) {
    diffs <- ( (arr[i,]) - theor)^2
    u[i] <- sum( diffs)
  }
  print(u)
  signif( (rank(-u)[1])/n.s, 5)
}

lplot <- function(arr, ci=NULL, r=NULL, ylab, do.krank=FALSE, ...) {
  ## Plot a spatial distribution (K, F, G).
  ## Real data is shown in red; simulation envelope in black.
  if (is.null(r)) {
    r <- colnames(arr)
  }
  if (missing (ylab)) {
    ylab <- attributes(arr)$name
    if (is.null(ylab))
      ylab <- deparse(substitute(arr))
  }
  ## are we plotting a g function
  gfunction <- (max(arr[1,]) < 1.1)
  if (gfunction) {
    ylim <- c(0,1)
  } else {
    ylim <- c(0, 150)
  }

  if (do.krank) {
    pval <- kranking(arr)
  } else {
    pval <- ranking(arr)
  }
  plot(r, arr[1,], col='black', type='l', bty='n',
       ##main=ranking(arr),
       ylim=ylim,
       ylab=ylab, xlab='')
  title(xlab=expression(paste("distance (", mu, "m)")), mgp=c(2,1,0))
  ##text(0, 140, paste('p = ', pval), adj=0)
  mtext(paste('p = ', pval), adj=0.1, side=3, line=-1, cex=0.8)
  ##title(main= paste("p =", ranking(arr)), line=-1)
  if (is.null(ci)) {
    ## if no CI given, just plot envelope.
    min.line <- apply(arr[-1,], 2, min)
    max.line <- apply(arr[-1,], 2, max)
    lines(r, min.line, lty=1);   lines(r, max.line, lty=1)
  } else {
    ## We have the CI data, so plot it.
    lines(r, ci[1,], lty=2)
    lines(r, ci[2,], lty=2)
  }
}

w81s1.plot <- function(pdf=TRUE){
  if (pdf) {
    pdf(file="w81s1_gof2t.pdf",
        onefile=FALSE, width=7, height=9)
  }
  par(mar=c(3,4, 1, 1), mfrow=c(4,2), bty='n', las=1,oma=c(1,0,0,0))
  lplot(w81s1.arr$get()$l1,  w81s1.ci$l1,
        ylab=expression(L[1]))
  lplot(w81s1.arr$get()$l2,  w81s1.ci$l2,
        ylab=expression(L[2]))
  lplot(w81s1.arr$get()$l0,  w81s1.ci$l0,
        ylab=expression(L[1+2]))
  lplot(w81s1.arr$get()$l12, w81s1.ci$l12,
        ylab=expression(L[12]))
  lplot(w81s1.arr$get()$g1,  w81s1.ci$g1,  ylab=expression(G[1]))
  lplot(w81s1.arr$get()$g2,  w81s1.ci$g2,  ylab=expression(G[2]))
  plot.spat.opp2(w81s1.arr$get()$opp, cex=0.1, real.col='black')
  plot.spat.ri3.v2(w81s1.arr$get()$ri3, cex=0.1, real.col='black')
  ##mtext('W81s1 29 Oct 2004; bivariate pipp theta1=theta2=(45, 41.53, 2.66), d12=18', side=1, outer=T)
  if (pdf) {dev.off()}
}

m623.plot <- function(pdf=TRUE) {
  if(pdf) {
    pdf(file="m623_gof2t.pdf",
        onefile=FALSE, width=7, height=9)
  }
  par(mar=c(3,4, 1, 1), mfrow=c(4,2), bty='n', las=1,oma=c(1,0,0,0))
  lplot(m623.arr$get()$l1,  m623.ci$l1,   ylab=expression(L[1]))
  lplot(m623.arr$get()$l2,  m623.ci$l2,   ylab=expression(L[2]))
  lplot(m623.arr$get()$l0,  m623.ci$l0,   ylab=expression(L[1+2]))
  lplot(m623.arr$get()$l12, m623.ci$l12, ylab=expression(L[12]))
  lplot(m623.arr$get()$g1,   m623.ci$g1,  ylab=expression(G[1]))
  lplot(m623.arr$get()$g2,   m623.ci$g2,  ylab=expression(G[2]))
  plot.spat.opp2(m623.arr$get()$opp, cex=.1, real.col='black')
  plot.spat.ri3.v2(m623.arr$get()$ri3, cex=.1, real.col='black')
  ##mtext('W81s1 29 Oct 2004; bivariate pipp theta1=theta2=(45, 41.53, 2.66), d12=18', side=1, outer=T)
  if (pdf) {dev.off()}
}

m623.handplot <- function() {
  postscript(file="m623_handgof.ps", horiz=F,
             onefile=FALSE, width=7, height=7)
  par(mar=c(3,4, 1, 1), mfrow=c(2,2), bty='n', las=1,oma=c(1,0,0,0))
  lplot(m623.arr$get()$l1,  m623.ci$l1,   ylab=expression(L[1]))
  lplot(m623.arr$get()$l12, m623.ci$l12, ylab=expression(L[12]))
  lplot(m623.arr$get()$g1,   m623.ci$g1,  ylab=expression(G[1]))
  plot.spat.ri3.v2(m623.arr$get()$ri3, cex=.1, real.col='black')
  dev.off()
}

plot.spat.opp2 <- function(arr,cex=0.5, real.col='black', ...) {
  ## Plot the fraction of opposites

  opp <- arr[-1,]
  boxplot( list(opp[,1], opp[,2], opp[,3], opp[,4]), xaxt ='n',
          ylab="fraction opposite type",
          pch=19, cex=0.2, pars=list(medlwd=1))

##   stripchart(list(arr[-1,1], arr[-1,2], arr[-1,3], arr[-1,4]),
##              method="jitter", pch=19, vertical=TRUE,
##              ylim=c(min(arr), 1), 
##              ##group.names=c(expression(1^{st}), "2", "3", "all"),
##              group.names=rep("", 4),
##              ylab="fraction opposite type", cex=cex, ...)

  axis(1, at=1:4, labels=c(expression(1^{st}),
                    expression(2^{nd}),
                    expression(3^{rd}),
                    "all"))
  dx <- 0.6; i <- 1:4
  ##segments(i-dx, arr[1,], i+dx, arr[1,], lwd=0.6, col=real.col)
  points(i, arr[1,], pch=18, cex=2)

  ##median.sim <- apply(arr[-1,], 2, median)
  ##segments(i-dx, median.sim, i+dx, median.sim, lwd=0.6, lty=2)

}


plot.spat.ri3.v2 <- function(ri3, cex=0.5, ylim=range(ri3),
                          real.col = 'red', ...) {
  ## Plot the regularity indexes.
  res <- list(on=ri3[-1,1], of=ri3[-1,2],on.off=ri3[-1,3])
  boxplot(res, ylab="regularity index",
          names = c("ON", "OFF", "ON+OFF"),
          pch=19, cex=0.2, pars=list(medlwd=1))
##   stripchart(res, vert=T, pch=19, method="jitter",
##              cex=cex, ylim=ylim,
##              group.names=c("ON", "OFF", "ON+OFF"),
##              main="", ylab="regularity index")
  

  i <- 1:3; dx <- 0.3;
  ##segments(i-dx, ri3[1,], i+dx, ri3[1,], lwd=0.6, col=real.col)
  points(i, ri3[1,], pch=18, cex=2)

  ## median.sim <- apply(ri3[-1,], 2, median)
  ## segments(i-dx, median.sim, i+dx, median.sim, lwd=0.6, lty=2)
  ##legend(x=1, y=3.5, lty=c(1,2),
  ##       legend=c("experimental RI", "median RI of sims"))
  
}


h.nopar <- function(pts, win, rs, plot=FALSE) {
  ## Adapted from ~/mosaics/beta_rgc/pipp/plot_nonparh.R
  ##
  ## Compute the non-parametric estimate of h() for a dataset.
  ## Input:
  ## PTS - (N,2) matrix of data points.
  ## WIN - 4 element window.
  ## NTILES - 2-vector giving number of tiles in (x,y) dimension for
  ## quadrature.
  ## RS - vector of breakpoints where h() is estimated.
  ##
  ## Output: list(x,y) estimating the h() function.
    
  ppp <- as.ppp(pts, as.owin(win))
  qs <- quadscheme(ppp)
  x <- ppm(qs, ~1, PairPiece(r = rs), correction="isotropic")
  h <- summary(x)$interaction$printable

  if (any(is.na(h)))
    h[which(is.na(h))] <- 0
  
  ## return the list of (x,y) points.
  
  ## do we want to take the mid points of RS?  RS are the breaks.
  x <- apply(rbind(rs, c(0, rs[1:(length(rs)-1)])), 2, mean)
  list(x=x, y=h)
}

plot.hnopar <- function(arr, xs, p=0.05) {
  ## Plot the non-parametric estimate of h and confidence intervals.

  ci <- apply(arr[-1,], 2, quantile, probs=c(p/2, 1-(p/2)))
  plot(xs, arr[1,], type='l', col='red', main='',
       xlab=expression(paste("distance (", mu, "m)")),
       ylab="h",
       ylim=c(0,1.8))
  title(deparse(substitute(arr)))
  lines(xs, ci[1,]); lines(xs, ci[2,])
}
```

# Evaluation of mosaics

```{r}
vd.num <- 30
m623.bdpar <- list( steps=seq(from=0, to=150, length=50),
                   vd0.breaks=seq(from=0, to=12000, len=vd.num),
                   vd1.breaks=seq(from=0, to=24000, len=vd.num),
                   ds0.breaks=seq(from=0, to=200, len=vd.num),
                   ds1.breaks=seq(from=0, to=250, len=vd.num),
                   distribs=list(g0=1, g1=1,g2=1,
                     f0=1, f1=1,f2=1,
                     l0=1, l1=1, l2=1, l12=1,
                     ##ia0=1, ia1=1, ia2=1,
                     vd0=1, vd1=1, vd2=1,
                     ds0=1, ds1=1, ds2=1,
                     opp=1, ri3=1))


vd.num <- 30                            #given less than 100 real pts
w81s.bdpar <- list( steps=seq(from=1, to=150, length=100),
                   vd0.breaks=seq(from=0, to=12000, len=vd.num),
                   vd1.breaks=seq(from=0, to=20000, len=vd.num),
                   ds0.breaks=seq(from=0, to=200, len=vd.num),
                   ds1.breaks=seq(from=0, to=250, len=vd.num),
                   distribs=list(g0=1, g1=1,g2=1,
                     f0=1, f1=1,f2=1,
                     l0=1, l1=1, l2=1, l12=1,
                     ##ia0=1, ia1=1, ia2=1,
                     vd0=1, vd1=1, vd2=1,
                     ds0=1, ds1=1, ds2=1,
                     opp=1, ri3=1))


## Should we look at non-parametric estimates of h?
estimate.h <- FALSE

nreps <- 99

##Fri 03 Jun 2005: trying hand generated params.
try.hand <- FALSE
```

## Try M623 Goodness of fit.

```{r}

## Print parameter estimates of the four fields.
print(w81s1.on.par)
print(w81s1.of.par)
print(m623.of.par)
print(m623.on.par)


if (try.hand) {
  ##m623.on.par <- c(15, 75, 8)           #ok
  m623.on.par <- c(15, 75, 10)          #ok2
}

h11.x <- seq(from=0, to=250, by=5); h22.x <- h11.x
h11.y <- hpar(h11.x, m623.on.par)
h22.y <- hpar(h22.x, m623.of.par)


h12.x <- seq(from=19.5, to=20.5, by=.05)
h12.y <- ifelse(h12.x>20.0, 1, 0)

plot(h11.x, h11.y, ylim=c(0, 1.5), type='l', col='green')
lines(h22.x, h22.y, col='orangered')
lines(h12.x, h12.y, col='black')
legend(100, .2, c(expression(h[11]), expression(h[22]), expression(h[12])),
       text.col= c("green", "red", "black"))



m623.arr <- new.dist.arr( sjespatdists.biv(m623.on, m623.of, m623.w, "note",
                                           param=m623.bdpar), nreps)
n1 <- nrow(m623.on); n2 <- nrow(m623.of)


if (estimate.h) {
  m623.rs <- seq(from=10, to=200, by=10)
  m623.h.on <- m623.h.of <- matrix(0, nrow=1+nreps, ncol=length(m623.rs))
  h.on1 <- h.nopar(m623.on, m623.w, m623.rs)
  m623.h.on[1,] <- h.on1$y
  h.of1 <- h.nopar(m623.of, m623.w, m623.rs)
  m623.h.of[1,] <- h.of1$y
}

sim.on <- m623.on; sim.of <- m623.of
for (i in 1:nreps) {
  if ((i %%10) == 0)
    cat(paste("iteration", i,"\n"))
  nsweeps <- ifelse(i==1, 10, 1)
  d <- pipp2.lookup(w=m623.w,
                    pts1=sim.on, pts2=sim.of,
                    n1=n1, n2=n2,
                    h1=h11.y, d1=h11.x,
                    h2=h22.y, d2=h22.x,
                    h12=h12.y, d12=h12.x, tor=FALSE,
                    nsweeps=nsweeps, verbose=FALSE)

  simpts <- cbind(d$x, d$y)
  sim.on <- simpts[1:(d$n1),]; sim.of <- simpts[-(1:(d$n1)),]
  s <- sjespatdists.biv(sim.on, sim.of, m623.w, "note", param=m623.bdpar)
  m623.arr$set.row(s, i+1)

  if (estimate.h) {
    ## Check non-par estimates.
    h.on1 <- h.nopar(sim.on, m623.w, m623.rs)
    m623.h.on[i+1,] <- h.on1$y
    
    h.of1 <- h.nopar(sim.of, m623.w, m623.rs)
    m623.h.of[i+1,] <- h.of1$y
  }
}

if (estimate.h) {
  par(mfrow=c(1,2))
  plot.hnopar(m623.h.on, m623.rs)
  plot.hnopar(m623.h.of, m623.rs)
}

m623.sim.on <- sim.on; m623.sim.of <- sim.of
m623.ci <- make.ci(m623.arr, c("l1", "l2", "l0", "l12", "g1", "g2"))

##load("bdmin_w81_jul01.Rda")
save(m623.arr, sim.on, sim.of, m623.ci,file="bpipp_m623_date_1.Rda",
     compress=TRUE)
##save(m623.sim.on, m623.sim.of, w81s1.sim.on, w81s1.sim.of, file='bpipp_examples.Rda')

m623.plot()


## If just wanting to regenerate handplot, try following.
if (try.hand) {
  ## this one has the data to use.
  ##save.image("bpipp_hand_jun3.Rda", compress=TRUE)
  m623.handplot()
}

```

## Try W81s GOF
```{r}
h11.x <- seq(from=0, to=160, by=5); h22.x <- h11.x
h11.y <- hpar(h11.x, w81s1.on.par)
h22.y <- hpar(h22.x, w81s1.of.par)

##load("~/mosaics/beta_rgc/pipp/w81s1.nopar")
##h11.x <- w81s1.on.hnopar$x; h11.y <- w81s1.on.hnopar$y

h12.x <- seq(from=17.5, to=18.5, by=.05)
h12.y <- ifelse(h12.x>18.0, 1, 0)

par(mfrow=c(1,2))
plot(h11.x, h11.y, type='l', col='green',
     xlab="distance (um)",
     ylab="prob. of acceptance")
lines(h22.x, h22.y, col='orangered')
lines(h12.x, h12.y, col='black')
legend(100, .2, c(expression(h[11]), expression(h[22]), expression(h[12])),
       text.col= c("green", "red", "black"))



n1 <- nrow(w81s1.on); n2 <- nrow(w81s1.of)

w81s1.arr <- new.dist.arr( sjespatdists.biv(w81s1.on, w81s1.of, w81s1.w, "note",
                                           param=w81s.bdpar), nreps)

w81s1.rs <- seq(from=10, to=200, by=10)

if (estimate.h) {
  w81s1.h.on <- w81s1.h.of <- matrix(0, nrow=1+nreps, ncol=length(w81s1.rs))
  h.on1 <- h.nopar(w81s1.on, w81s1.w, w81s1.rs)
  w81s1.h.on[1,] <- h.on1$y
  h.of1 <- h.nopar(w81s1.of, w81s1.w, w81s1.rs)
  w81s1.h.of[1,] <- h.of1$y
}

sim.on <- w81s1.on; sim.of <- w81s1.of
for (i in 1:nreps) {
  if ((i %%10) == 0)
    cat(paste("iteration", i,"\n"))
  nsweeps <- ifelse(i==1, 10, 10)
  d <- pipp2.lookup(w=w81s1.w,
                    pts1=sim.on, pts2=sim.of,
                    n1=n1, n2=n2,
                    h1=h11.y, d1=h11.x,
                    h2=h22.y, d2=h22.x,
                    h12=h12.y, d12=h12.x, tor=FALSE,
                    nsweeps=nsweeps, verbose=FALSE)

  simpts <- cbind(d$x, d$y)
  sim.on <- simpts[1:(d$n1),]; sim.of <- simpts[-(1:(d$n1)),]
  s <- sjespatdists.biv(sim.on, sim.of, w81s1.w, "note", param=w81s.bdpar)
  w81s1.arr$set.row(s, i+1)

  if (estimate.h) {
    ## Check non-par estimates.
    h.on1 <- h.nopar(sim.on, w81s1.w, w81s1.rs)
    w81s1.h.on[i+1,] <- h.on1$y
    
    h.of1 <- h.nopar(sim.of, w81s1.w, w81s1.rs)
    w81s1.h.of[i+1,] <- h.of1$y
  }
}

if (estimate.h) {
  par(mfrow=c(1,2))
  plot.hnopar(w81s1.h.on, w81s1.rs)
  plot.hnopar(w81s1.h.of, w81s1.rs)
}


w81s1.sim.on <- sim.on; w81s1.sim.of <- sim.of

w81s1.ci <- make.ci(w81s1.arr, c("l1", "l2", "l0", "l12", "g1", "g2") )
w81s1.plot()

save(w81s1.arr, sim.on, sim.of, w81s1.ci,file="bpipp_w81s1_date_1.Rda",
     compress=TRUE)
```

# Output figures

```{r, fig.width=6, fit.height=6, fig.cap="W81S1 Goodness of fit (Figure 7 of paper)"}
w81s1.plot(FALSE)
```

```{r, fig.width=6, fit.height=6, fig.cap="M623 Goodness of fit (Figure 8 of paper)"}
m623.plot(FALSE)
```



# References

Eglen SJ, Diggle PJ, Troy JB (2005) Homotypic constraints dominate
positioning of on- and off-center beta retinal ganglion cells. Vis
Neurosci 22:859–871 Available at:
<https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1513157/>
